Transfer of Learned Opponent Models in Zero Sum Games

### 2021-07-05 Report/paper_draft_2021.Rmd
went through draft

abstract: |
  Human learning transfer abilities take advantage of important cognitive building blocks such as an abstract representation of concepts underlying tasks and 
  causal models of the environment. One way to build abstract representations of the environment when the task involves interactions with others is to build 
  a model of the opponent that may inform what actions they are likely to take next. In this study, we propose to explore opponent modelling and its transfer with the use of
  computer agents possessing human-like theory of mind abilities with limited degrees of iterated reasoning. Through two differentiated experiments, we find that participants can
  deviate from Nash equilibrium play and learn to adapt to the opponent strategy and exploit it. Moreover, we showed that participants transfer their learning to new games and that
  the transfer is moderated by the level of  sophistication of the opponent. Computational modelling shows that it is likely that players start each game using a model-based
  learning strategy that facilitates generalisation and opponent model transfer, but then switch to behaviour that is consistent with a  model-free learning strategy in the latter
  stages of the interaction.

@@

In this paper, we are specifically interested in the way in which people build and use models of their opponent to facilitate learning transfer, when engaged in situations involving an interaction with strategic considerations. Repeated games, in which players interact repeatedly with the same opponent and have the ability to learn about the opponent's strategies and preferences [@mertens1990repeated] are particularly adapted to this task. The early literature on learning transfer in repeated games has mostly focused on measuring the proportion of people who play normatively optimal (Nash Equilibria) or salient actions (e.g Risk Dominance) in later games, having had experience with a similar game environment previously [@ho1998iterated; @knez2000]. This doesn't allow for the possibility of learning about the opponent's strategy and potentially exploiting it.

When studies have specifically explored this aspect, they have used computer opponents that were generally programmed not to change their strategies over the course of the task, allowing better experimental control. However, they have mostly looked at the ability of players to detect and exploit action-based learning rules [@spiliopoulos2013strategic; @shachat2004we]. The strategies implemented by the computer opponents had a style of play that was not "human-like" in the sense that humans are not very good at playing specific mixed strategies with precision, or at detecting patterns from long sequences of past play. Thus, in this study, we aim to explore opponent modelling and its transfer with the use of computer agents endowed with human-like limited degrees of iterated reasoning. The agents are either a level-1 or level-2 player, mimicking ``I know that you know that I know'' type reasoning, and the limited recursion depth they exhibit [@goodie2012levels]. A level 1 player adapts their play to what they believe their opponent will play, without considering what their opponent might believe they will play. A level 2 player, on the other hand, takes their opponent's belief about their actions into account, assuming they face a level 1 player, and choosing actions to beat the actions of that player. The choice of this type of strategy is also motivated by evidence that humans strategically use information from last round play of their opponents in zero sum games [@batzilis; @wang2014social].

@@

@@

### Tasks
In the first experiment, participants played the three games against their computer opponent. These games were rock-paper-scissors, fire-water-grass, and the numbers game. A typical rock-paper-scissors game (hereafter RPS) is a 3x3 zero sum game, with a cyclical hierarchy between the two player's actions: rock blunts scissors, paper wraps rock, and scissors cut paper. If one player chooses an action which dominates their opponent's action, the player wins (receives a reward of 1) and the other player loses (receives a reward of -1). Otherwise it is a draw and both players receive a reward of 0. RPS has a unique mixed-strategy Nash equilibrium, which consists of each player in each round randomly selecting from the three options with uniform probability. The Fire-Water-Grass (FWG) game is identical to RPS in all but action labels: Fire burns grass, water extinguishes fire, and grass absorbs water. We use this game as we are interested in whether learning is transferred in a fundamentally similar game where the only difference is in the description of the possible actions. This should make it relatively easy to generalize knowledge of the opponent's strategy, provided this knowledge is on a sufficiently abstract level, such as knowing the opponent is a level 1 or 2 player. Crucially, learning simple contingencies such as ``If I played Rock on the previous round, playing Scissors next will likely result in a win'', as might be learned by a simple reinforcement learning algorithm, will not be able to generalize to such a game, as these contingencies are tied to the labels of the actions. The numbers game is a generalization of RPS. In the variant we use, 2 participants concurrently pick a number between 1 and 5. To win in this game, a participant needs to pick a number exactly 1 higher than the number chosen by their opponent. For example, if a participant thinks their opponent will pick 3, they ought to choose 4 to win the round. To make the strategies cyclical as in RPS, the game stipulates that the lowest number (1) beats the highest number (5), so if the participant thinks the opponent will play 5, then the winning choice is to pick 1. This game has a structure similar to RPS in which every action is dominated by exactly one other action. All other possible combinations of choices are considered ties. Similar to RPS and FWG, the mixed-strategy Nash equilibrium is to play each action with equal probability in a random way.

The computer opponent was programmed to use either a level-1 or level-2 strategy in all the games. A level 1 player is defined as a player who best responds to a level 0 player. A level 0 player plays in a non-strategic way and does not consider their opponent's actions. Here, we assume a level 0 player simply repeats their previous action. There are other ways to define a level 0 player. For instance, as repeating their action if it resulted in a win, and choosing randomly from the remaining actions otherwise. As a best response to a random action is itself a random action, defining a level 0 player in such a way would make a level 1 opponent's strategy much harder to discern. Because we are mainly interested in generalization of knowledge of an opponent's strategy to other games, which rests on good knowledge of this strategy, we opted for this more deterministic formulation of a level 0 player (whilst also introducing some randomness in the computer opponent's play). A level-2 computer opponent, will assume in turn that the participant is a level-1 opponent, playing according to the strategy just described. We also introduced some noise over the actions of computer opponents making them play randomly in 10\% of all trials.

\begin{table}[bth!]
\centering
\begin{tabular}{||c c c c||} 
 \hline
 Human last & Agent last & level-1 Agent & level-2 Agent\\ [0.5ex] 
 \hline\hline
 Paper & Rock & Scissors & Scissors \\ 
 Scissors  & Scissors & Rock & Paper \\
 Rock & Paper & Paper & Rock \\
 ... & ... & ... & ... \\ [1ex] 
 \hline
\end{tabular}
\caption{Example of how a level-1 and level-2 computer agent plays in response to actions taken in the previous round.}
\label{table:1}
\end{table}

@@

Participants were informed they would play three different games against the same computer opponent. Participants were told that the opponent cannot cheat and will choose its actions simultaneously without knowledge of the participant's choice. After providing informed consent and reading the instructions, participants answered a number of comprehension questions. They then played the three games against their opponent in the order RPS, FGW, and NUMBERS. A total of 50 rounds of each game was played with the player's score displayed at the end of each game. The score was calculated as the number of wins minus the number of losses. Ties did not affect the score. In order to incentivise the participants to maximise the number of wins against the opponents, players were paid a bonus at the end of the experiment that was proportional to their final score. Each point is worth £0.02. An example of the interface for the RPS game is provided in Figure\ \@ref(fig:feedback-rps-exp2). After playing all the games, participants were asked questions about their beliefs about the computer opponent, related to whether they think they have learned their strategy and how hard they found playing against that particular opponent. They were then debriefed and thanked for their participation.

@@

```{r feedback-rps-exp2, fig.cap = "Screenshot of the feedback at the end of a round of Rock-Paper-Scissors", fig.align='center'}

@@

```{r exp1-avg-scores-game, fig.cap = "Boxplot of scores per game across conditions", fig.align="center"}
# Plot average  scores per game 
ggboxplot(dat3, x = "game", y = "game_score",group = 1, add =c("mean_ci"), color="game", palette = c("#00AFBB", "#E7B800", "#FC4E07"), order = c("rps", "fwg","numbers"), ylab = "Score", xlab = "Game", legend.title = "Game") + geom_jitter( aes(color = game), width = 0.2, alpha = 0.3 )

```
On average, participants obtained the lowest score in the RPS game ($M = 0.289$, $SD = 0.348$), followed by NUMBERS ($M = 0.31$, $SD = 0.347$). Participants' performance was highest in the FWG game ($M = 0.454$, $SD = 0.354$). Scores in each game were significantly different from 0, the expected score of random play (RPS: $t(51) = 7.26$, $p < .001$;  FWG:  $t(51) = 10.04$ , $p < .001$; NUMBERS: $t(51) = 7.17$, $p < .001$). To assess learning within and between games, we used a 2 (condition: level-1, level-2) by 3 (game: RPS, FWG, NUMBERS) by 2 (block: first half, second half) repeated-measures ANOVA, with the first factor varying between participants. This showed a main effect of Game ($F(2,100) = 8.54$, $\eta^{2}  = 0.05$,  $p < .001$), indicating that average scores varied significantly over the games. Post-hoc pairwise comparisons showed that performance in the FWG game was significantly higher than in the RPS game ($t(100) =3.78$, $p < .001$) and the NUMBERS game ($t(100) = 3.32$ , $p = .002$). The score in RPS was not significantly different from the score in NUMBERS ($t(100) = 0.45$ , $p = .65$). The main effect of Block ($F(1,50) = 22.51$ ,  $\eta^{2} = 0.03$, $p < .001$) shows that the score in the first half of each game ($M = 0.29$) was significantly lower than in the second half ($M = 0.40$), which indicates within-game learning. The main effect of Condition ($F(1,50) = 5.44$, $\eta^{2} = 0.05$, $p = .024$) indicates that scores were higher against the level-1 player ($M = 0.43$) than against the level-2 player ($M = 0.27$). Thus, it appears that is was harder for participants to exploit the strategy of the more sophisticated level-2 opponent than the comparatively less sophisticated level-1 opponent.

@@

```{r exp1-3factor-plot, include= FALSE, fig.cap = "Boxplots of scores by game, block and condition", fig.align = "center"}

# Plot results by game and block (for all 3 games, learning is happening) 
p <- ggboxplot(dat2, x = "game", y = "block_score", palette = c("#00AFBB", "#E7B800"),order = c("rps", "fwg","numbers"), fill="block",ylab = "Percentage score", xlab = "Games")

p + facet_grid(. ~ condition)

@@

As a measure for learning transfer, we focus on participants’ scores in the initial 5 rounds after the first round (rounds 2-6) of each game (see Figure \ref{fig:exp1-score-by-opp}). We exclude the very first round as the computer opponent plays randomly here and there is no opportunity yet for the human player to exploit their opponent's strategy. Players with no knowledge of their opponent's strategy are expected to perform at chance level in these early rounds. Positive scores in rounds 2-6 reflect generalization of prior experience. The FWG early score score is significantly higher than 0 ($t(148.85) = 4.584$, $p < .001$). This is also the case for the NUMBERS game ($t(148.85) = 3.00$, $p = .009$). We did not expect positive scores for the RPS game, as it was the first game played and there was no opportunity for learning about the opponent's strategy. Scores in this game was indeed not significantly different from 0 ($t(148.85) = 1.04$ , $p = .89$).

@@

(ref:figure4-caption) Mean and standard error of early scores by game and opponent
```{r exp1-early-score-by-opp, fig.cap = "(ref:figure4-caption)", fig.align ='center'}

p2 <- ggerrorplot(exp1_dat2_6, x = "game.f", y = "early_score", group = 1, color="game.f", desc_stat = "mean_se",palette = c("#00AFBB", "#E7B800", "#FC4E07"), order = c("RPS", "FWG","Numbers"), ylab = "Early rounds score", xlab = "Games")

p2 + facet_grid(. ~ condition)

@@

Next, we explore whether learning transfer is moderated by the type of opponent and game similarity. We expected better transfer between more similar games (i.e. better transfer from RPS to FWG than from RPS/FWG to NUMBERS), and worse transfer for the more sophisticated level 2 agent. Figure~\ref{fig:exp1-early-score-by-opp} indicates that the pattern over the games is indeed dissimilar between level-1 and level-2 players. To explore this, we used a 2 (condition: level-1, level-2) by 3 (game: RPS, FWG, NUMBERS) repeated measures ANOVA with the first factor varying between participants. There was a main effect of Game ($F(2,92) = 3.35$, $\eta^{2}  = 0.04$,  $p < .04$). We then run statistical tests on early round scores by game and opponent against the null hypothesis of 0 (no transfer). For level-1 facing players, there is evidence of learning transfer from RPS to both FWG ( $t(150) = 3.96$, $p < .001$) and NUMBERS ($t(150) = 3.74$, $p < .001$). For level-2 facing players, there is evidence for transfer from RPS to the similar game FWG, albeit scores are lower than for level-1 player ($t(150) = 2.48$, $p = .01$) but not to the dissimilar game of NUMBERS.

### Experiment 1 discussion:

These results indicate that learning transfer to the more dissimilar game (NUMBERS) we found earlier is exclusively driven by level-1 facing players, as average early round scores in the NUMBERS game of level-2 facing players are close to 0. Therefore, both participants facing level-1 and level-2 agents can transfer learning to the similar FWG game, but only those facing the less sophisticated opponent are able to generalise to the less similar NUMBERS game.

@@

We ran a second experiment with various differentiated features to improve the opportunity to measure learning transfer. Instead of making participants face either the level-1 or level-2 player throughout, we made them face both opponents sequentially. Because there were two distinct opponents, requiring potentially holding two opponent models in memory, we also made it easier to recall the results of past rounds by providing participants with the opportunity to see the history of the game since the beginning of each interaction. Figure\ \@ref(fig:feedback-rps-exp2) shows an example of showing interaction history in the RPS game. Finally, we changed the third game to a penalty shootout game, whith participants aiming to score a goal and opponents playing the role of goal keepers.  Whilst this game has the same number of actions as the first two, it is strategically dissimilar. Unlike the third game in the first experiment, it did not have a cyclical hierarchy between actions, making it harder to win by just following simple heuristics leveraging this cyclicality. If we see evidence for differential play against opponents, it would show participants adapting their strategies to the opponent they are facing, which is indicative of opponent modelling.

@@

A total of 48 participants (21 females, 28 males, 1 unknown) used the Prolific Academic platform to participate in the experiment. This was a new set of participants unrelated to those taking part in Experiment 1. The average age of players was 30.2 years, and the mean duration to complete the task was 39 minutes. Participants were incentivised using a two-tier payment mechanism: a fixed fee of £2.5 for completing the experiment plus a performance linked bonus that averaged £1.32.

@@

knitr::include_graphics("images/shootout.png")
```

@@

>>>>

abstract: |
  Human learning transfer abilities take advantage of important cognitive building blocks such as an abstract representation of concepts underlying tasks and 
  causal models of the environment. One way to build abstract representations of the environment when the task involves interactions with others is to build 
  a model of the opponent that may inform what actions they are likely to take next. In this study, we explore opponent modelling and its transfer with the use of
  computer agents with human-like limited degrees of iterated reasoning. In two experiments, we find that participants
  deviate from Nash equilibrium play and learn to adapt to the opponent's strategy and exploit it. Moreover, we show that participants transfer their learning to new games and that
  this transfer is moderated by the level of sophistication of the opponent. Computational modelling shows that players start each game with a model-based
  learning strategy that facilitates between-game transfer of the opponent's strategy, but then switch to behaviour that is consistent with a model-free learning strategy in the latter stages of the interaction.

@@

A model of the opponent can help increase performance in a particular game, but learning more general characteristics of an opponent may also help increase performance in other games. In this paper, we are specifically interested in the latter: How do people build and use models of their opponent to facilitate learning transfer? <!-- when engaged in situations involving an interaction with strategic considerations.--> Repeated games, in which players interact repeatedly with the same opponent and have the ability to learn about the opponent's strategies and preferences [@mertens1990repeated] are particularly useful to address this question. The early literature on learning transfer in repeated games has mostly focused on the proportion of people who play normatively optimal (e.g. Nash equilibrium play) or use salient actions (e.g. risk dominance) in later games, having had experience with a similar game environment previously [@ho1998iterated; @knez2000]. As is well-known, a Nash equilibrium means that all players in a game play such that no-one can unilaterally improve their performance by deviating from their strategy. When playing against an opponent with a Nash-optimal strategy, I can do no better than play according to the Nash-equilibrium strategy as well. However, when faced with a player who deviates from the Nash-optimal strategy, it may be better for me to exploit this and also deviate from this strategy, increasing my performance beyond what is expected at a Nash equilibrium. Of course, this comes with some risk, as my own deviation from Nash-optimal play may leave myself open to similar exploitation.

<!-- his doesn't allow for the possibility of learning about the opponent's strategy and potentially exploiting it. !-->

Studies that focused on whether people can learn to exploit deviations from Nash-equilibrium play have mostly looked at the ability of players to detect and exploit action-based learning rules [@spiliopoulos2013strategic; @shachat2004we]. These studies used computer opponents with fixed strategies (no adaptation to their human opponent) consisting of playing each action with a fixed probability (a mixed strategy). Findings showed .... However, the use of mixed strategies may have limited people's ability to form accurate opponent models. Rather than thinking of their opponents and drawing a random action from a (non-uniform) probability distribution, people may rather think of their opponents as applying a form of iterated reasoning to determine their next action. The type of reasoning we refer to takes the form of ``I believe that you believe that I believe ...'' and has been shown to correspond to people's casual introspection (REFERENCE). It also underlies successful models in behavioural economics, such as Level-$k$ and Cognitive Hierarchy models, which posit that people assume their opponent has a limited level of iterative reasoning, with the player best responding to their (modelled) opponent.

In Level-$k$ theory, a level-0 player uses a fixed strategy without explicitly considering the strategy of their opponent. A level-1 player assumes their opponent is a level-0 player, and chooses actions to best respond to the strategy of their opponent, without considering what their opponent might believe that they will play. A level-2 player, on the other hand, takes their opponent's belief about their actions into account, assuming they face a level-1 player, and choosing actions to best respond to the actions of that player. Cognitive Hierarchy theory works on similar principles, but rather than assuming an opponent always adopts a particular level-$k$ strategy, they assume their opponent adopts each pf the level-$k$ strategies with a particular probability (i.e., the strategy is a mixture over pure level-$k$ strategies).

Iterative reasoniing strategies can explain non-equilibrium play in a range of games, such as ... In the present study, we endow our computer opponents with limited ability for iterative reasoning and assess whether (1) human players adapt their strategy to exploit this limited reasoning of their opponent, and (2) whether they are able to generalize a learned opponent model to other games. In two experiments, participants face the same opponent (Experiment 1) or same two opponents (Experiment 2) in three consecutive games: the well-known Rock-Paper-Scissors game, a structurally similar Fire-Water-Grass game, and a less similar Numbers (Experiment 1) or Shootout (Experiment 2) game. To foreshadow our results, we find evidence that participants transfer the learned strategy of their opponent to other games, but that this transfer is moderated by the sophistication of the opponent and the similarity of the games. Moreover, using computational modelling, we find evidence that participants switch from relying on an opponent model in the early stages of the games, to a morehabiltual and cognitively less demanding strategy in the later stages of the games.

<!--
with that were generally programmed not to change their strategies over the course of the task, allowing better experimental control. However, they have mostly looked at the ability of players to detect and exploit action-based learning rules [@spiliopoulos2013strategic; @shachat2004we]. The strategies implemented by the computer opponents had a style of play that was not "human-like" in the sense that humans are not very good at playing specific mixed strategies with precision, or at detecting patterns from long sequences of past play. Thus, in this study, we aim to explore opponent modelling and its transfer with the use of computer agents endowed with human-like limited degrees of iterated reasoning. The agents are either a level-1 or level-2 player, mimicking ``I know that you know that I know'' type reasoning, and the limited recursion depth they exhibit [@goodie2012levels]. A level 1 player adapts their play to what they believe their opponent will play, without considering what their opponent might believe they will play. A level 2 player, on the other hand, takes their opponent's belief about their actions into account, assuming they face a level 1 player, and choosing actions to beat the actions of that player. The choice of this type of strategy is also motivated by evidence that humans strategically use information from last round play of their opponents in zero sum games [@batzilis; @wang2014social].

@@

In the first experiment, we .... [intro to first exp]

@@

### Tasks
In the first experiment, participants played the three games against their computer opponent. These games were rock-paper-scissors, fire-water-grass, and the numbers game. A typical rock-paper-scissors game (hereafter RPS) is a 3x3 zero sum game, with a cyclical hierarchy between the two player's actions: rock blunts scissors, paper wraps rock, and scissors cut paper. If one player chooses an action which dominates their opponent's action, the player wins (receives a reward of 1) and the other player loses (receives a reward of -1). Otherwise it is a draw and both players receive a reward of 0. RPS has a unique mixed-strategy Nash equilibrium, which consists of each player in each round randomly selecting from the three options with uniform probability. The Fire-Water-Grass (FWG) game is identical to RPS in all but action labels: Fire burns grass, water extinguishes fire, and grass absorbs water. We use this game as we are interested in whether learning is transferred in a fundamentally similar game where the only difference is in the name of the possible actions. This should make it relatively easy to generalize knowledge of the opponent's strategy, provided this knowledge is on a sufficiently abstract level, such as knowing the opponent is a level-1 or level-2 player. Crucially, learning simple contingencies such as ``If I played Rock on the previous round, playing Scissors next will likely result in a win'', <!-- as might be learned by a simple reinforcement learning algorithm,--> is not generalizable to this similar game, as these contingencies are tied to the labels of the actions. The Numbers game is a generalization of RPS. In the variant we use, 2 participants concurrently pick a number between 1 and 5. To win in this game, a participant needs to pick a number exactly 1 higher than the number chosen by their opponent. For example, if a participant thinks their opponent will pick 3, they ought to choose 4 to win the round. To make the strategies cyclical as in RPS, the game stipulates that the lowest number (1) beats the highest number (5), so if the participant thinks the opponent will play 5, then the winning choice is to pick 1. This game has a structure similar to RPS in which every action is dominated by exactly one other action. All other possible combinations of choices are considered ties. Similar to RPS and FWG, the mixed-strategy Nash equilibrium is to randomly play each action with equal probability.

The computer opponent was programmed to use either a level-1 or level-2 strategy in all the games. A level-1 player is defined as a player who best responds to a level-0 player. A level-0 player plays in a non-strategic way and does not consider their opponent's actions. Here, we assume a level-0 player simply repeats their previous action. There are other ways to define a level-0 player. For instance, as repeating their action if it resulted in a win and choosing randomly from the remaining actions otherwise, or choosing randomly from all actions. As a best response to a uniformly random action is itself a random action, defining a level-0 player in such a way would make a level-1 opponent's strategy much harder to discern. Because we are mainly interested in generalization of knowledge of an opponent's strategy to other games, which rests on good knowledge of this strategy, we opted for this more deterministic formulation of a level-0 player (whilst also introducing some randomness in the computer opponent's play). A level-2 computer opponent, will assume in turn that the participant is a level-1 opponent, playing according to the strategy just described. We also introduced some noise over the actions of computer opponents making them play randomly in 10\% of all trials. Note that at all levels, the strategies are contingent on the actions taken in the previous round. The choice of this type of strategy is consistent with evidence that humans strategically use information from last round play of their opponents in zero sum games [@batzilis; @wang2014social].

\begin{table}[bth!]
\centering
\begin{tabular}{||c c | c c||} 
 \hline
 Human action $t-1$ & Computer action $t-1$  & Computer level-1 action $t$ & Computer level-2 action $t$ \\ [0.5ex] 
 \hline\hline
 Paper & Rock & Scissors & Scissors \\ 
 Scissors  & Scissors & Rock & Paper \\
 Rock & Paper & Paper & Rock \\
 ... & ... & ... & ... \\ [1ex] 
 \hline
\end{tabular}
\caption{Example of how a level-1 and level-2 computer agent plays in response to actions taken in the previous round.}
\label{table:1}
\end{table}

@@

Participants were informed they would play three different games against the same computer opponent. Participants were told that the opponent cannot cheat and will choose its actions simultaneously with them, without prior knowledge of the participant's choice. After providing informed consent and reading the instructions, participants answered a number of comprehension questions. They then played the three games against their opponent in the order RPS, FGW, and Numbers. An example of the interface for the RPS game is provided in Figure\ \@ref(fig:feedback-rps-exp2). On each round, tHe human player chooses an action, and after a random delay, is shown the action chosen by the computer opponent, and the outcome of that round. A history of the chosen actions over all rounds is available in the centre panel. A total of 50 rounds of each game was played with the player's score displayed at the end of each game. The score was calculated as the number of wins minus the number of losses. Ties did not affect the score. In order to incentivise the participants to maximise the number of wins against the opponents, players were paid a bonus at the end of the experiment that was proportional to their final score. Each point is worth £0.02. After playing all the games, participants were asked questions about their beliefs about the computer opponent, related to whether they thought they learned their opponent's strategy, and how difficult they found playing against their particular opponent. They were then debriefed and thanked for their participation.

@@

```{r feedback-rps-exp2, fig.cap = "Screenshot of the Rock-Paper-Scissors game in Experiment 2. Shown here is the feedback stage, after both the human (left) and computer (right) players have chosen their action. The interface was similar in Experiment 1, but excluded the history of game play in the center panel.", fig.align='center'}

@@

```{r exp1-avg-scores-game, fig.cap = "Performance per game and block across conditions in Experiment 1. Points are scores of individual participants and boxes reflect the 95\\% confidence intervals of the mean (center line equals the mean).", fig.align="center", fig.width=8, fig.height = 4}
# Plot average  scores per game

dat2 %>%
  mutate(game = factor(game)) %>%
  mutate(game = forcats::fct_recode(game, RPS = "rps", FWG = "fwg", Numbers = "numbers")) %>%
  mutate(game = factor(game, levels = c("RPS", "FWG", "Numbers"))) %>%
  mutate(condition = factor(condition, levels = c("Level1", "Level2"))) %>%
  mutate(condition = forcats::fct_recode(condition, "Level 1" = "Level1", "Level 2" = "Level2")) %>%
  ggplot(aes(x = game, y = block_score, colour = block)) + stat_summary(fun.data = "mean_ci", position = position_dodge2(width=.8), geom = "crossbar")  + geom_point(position=position_jitterdodge(), alpha=.4) + ylab("Score") + xlab("Game") + facet_grid(. ~ condition) + theme_apa() + geom_hline(yintercept = 0, lty=2)

#+ geom_boxplot(alpha=.4) + geom_point(position=position_jitterdodge(), alpha=.4) + ylab("Score") + xlab("Game") + facet_grid(. ~ condition) + theme_apa() + geom_hline(yintercept = 0, lty=2)

#geom_jitter(aes(colour=block, fill=block), width=.2) + facet_grid(. ~ condition)

# ggboxplot(x = "game", y = "block_score", colour="block", fill="block", add =c("mean_ci"), palette = c("#00AFBB", "#E7B800", "#FC4E07"), order = c("RPS", "FWG","NUMBERS"), ylab = "Score", xlab = "Game", legend.title = "Block") + geom_jitter(aes(color = block, fill = block), width = 0.2, alpha = 0.3 ) + facet_grid(. ~ condition)

```
On average, participants obtained the lowest score in the RPS game ($M = 0.289$, $SD = 0.348$), followed by Numbers ($M = 0.31$, $SD = 0.347$). Participants' performance was highest in the FWG game ($M = 0.454$, $SD = 0.354$). Scores in each game were significantly different from 0, the expected score of uniformly random play (RPS: $t(51) = 7.26$, $p < .001$;  FWG:  $t(51) = 10.04$ , $p < .001$; Numbers: $t(51) = 7.17$, $p < .001$). As uniformly random play is the Nash equilibrium, this indicates successful deviation from a Nash-optimal strategy.

To assess learning within and between games, we used a 2 (condition: Level 1, Level 2) by 3 (game: RPS, FWG, Numbers) by 2 (block: first half, second half) repeated-measures ANOVA, with the first factor varying between participants. This showed a main effect of Game ($F(2,100) = 8.54$, $\eta^{2}  = 0.05$,  $p < .001$), indicating that average scores varied significantly over the games. Post-hoc pairwise comparisons showed that performance in the FWG game was significantly higher than in the RPS game ($t(100) =3.78$, $p < .001$) and the Numbers game ($t(100) = 3.32$ , $p = .002$). The score in RPS was not significantly different from the score in Numbers ($t(100) = 0.45$ , $p = .65$). The main effect of Block ($F(1,50) = 22.51$ ,  $\eta^{2} = 0.03$, $p < .001$) shows that the score in the first half of each game ($M = 0.29$) was significantly lower than in the second half ($M = 0.40$), which indicates within-game learning. The main effect of Condition ($F(1,50) = 5.44$, $\eta^{2} = 0.05$, $p = .024$) indicates that scores were higher against the level-1 player ($M = 0.43$) than against the level-2 player ($M = 0.27$). Thus, it appears that is was harder for participants to exploit the strategy of the more sophisticated level-2 opponent than the comparatively less sophisticated level-1 opponent. **MS: Need to check test results and report GG degrees of freedom, etc.**

@@

```{r exp1-3factor-plot, include= FALSE, fig.cap = "Performance in early rounds (2-6) per game and block across conditions in Experiment 1. Points are scores of individual participants and boxes reflect the 95\\% confidence intervals of the mean (center line equals the mean).", fig.align = "center", fig.width=8, fig.height=4}

# Plot results by game and block (for all 3 games, learning is happening) 
dat2 %>%
  mutate(game = forcats::fct_recode(as.factor(game), RPS = "rps", FWG = "fwg", Numbers = "numbers")) %>%
  mutate(condition = factor(condition, levels = c("Level1", "Level2"))) %>%
  mutate(condition = forcats::fct_recode(condition, "Level 1" = "Level1", "Level 2" = "Level2")) %>%
  ggboxplot(x = "game", y = "block_score", palette = c("#00AFBB", "#E7B800"),order = c("RPS", "FWG","Numbers"), fill="block", ylab = "Percentage score", xlab = "Games") +  facet_grid(. ~ condition)

```{r ANOVAexp1, include = FALSE, warning=FALSE}

afex::aov_car(
  block_score ~ game.f*condition*block + Error(human_id/(game.f*block))
  , data=dat2
  , type = 3)

#apa_lm1 <- apa_print(exp1_early_anova)

#means.int2 <- lsmeans(exp1_early_anova, specs = c("game.f","condition"))
#trans26 <- summary(means.int2, infer = c(TRUE,TRUE),level = .95, adjust = "holm",ref=c("FWG","Numbers"))

@@

As a measure for learning transfer, we focus on participants’ scores in the initial 5 rounds after the first round (rounds 2-6) of each game (see Figure \ref{fig:exp1-early-score-by-opp}). We exclude the very first round as the computer opponent plays randomly here and there is no opportunity yet for the human player to exploit their opponent's strategy. Players with no knowledge of their opponent's strategy are expected to perform at chance level in these early rounds. Positive scores in rounds 2-6 reflect generalization of prior experience. The FWG early score score is significantly higher than 0 ($t(148.85) = 4.584$, $p < .001$). This is also the case for the Numbers game ($t(148.85) = 3.00$, $p = .009$). We did not expect positive scores for the RPS game, as it was the first game played and there was no opportunity for learning about the opponent's strategy. Scores in this game was indeed not significantly different from 0 ($t(148.85) = 1.04$ , $p = .89$).

@@

<!-- (ref:figure4-caption) Mean and standard error of early scores by game and opponent -->
```{r exp1-early-score-by-opp, fig.cap = "\\label{ref:figure4-caption}Performance in early rounds (2-6) per game and block across conditions in Experiment 1. Points are scores of individual participants and boxes reflect the 95\\% confidence intervals of the mean (center line equals the mean).", fig.align ='center', fig.width=8, fig.height=4}
exp1_dat2_6 %>%
  mutate(game = factor(game.f)) %>%
  mutate(game = forcats::fct_recode(game, Numbers = "Numbers")) %>%
  mutate(game = factor(game, levels = c("RPS", "FWG", "Numbers"))) %>%
  mutate(condition = factor(condition, levels = c("Level1", "Level2"))) %>%
  mutate(condition = forcats::fct_recode(condition, "Level 1" = "Level1", "Level 2" = "Level2")) %>%
  ggplot(aes(x = game, y = early_score, colour=game)) + stat_summary(fun.data = "mean_ci", geom = "crossbar")  + geom_jitter(alpha=.4) + ylab("Early score") + xlab("Game") + facet_grid(. ~ condition) + theme_apa() + geom_hline(yintercept = 0, lty=2) + scale_color_discrete() + theme(legend.position = "none")

#p2 <- ggerrorplot(exp1_dat2_6, x = "game.f", y = "early_score", group = 1, color="game.f", desc_stat = "mean_se",palette = c("#00AFBB", "#E7B800", "#FC4E07"), order = c("RPS", "FWG","Numbers"), ylab = "Early rounds score", xlab = "Games")

#p2 + facet_grid(. ~ condition)

@@

Next, we explore whether learning transfer is moderated by the type of opponent and game similarity. We expected better transfer between more similar games (i.e. better transfer from RPS to FWG than from RPS/FWG to Numbers), and worse transfer for the more sophisticated level 2 agent. Figure \ref{fig:exp1-early-score-by-opp} indicates that the pattern over the games is indeed dissimilar between level-1 and level-2 opponents. To explore this, we used a 2 (condition: Level 1, Level 2) by 3 (game: RPS, FWG, Numbers) repeated measures ANOVA with the first factor varying between participants. There was a main effect of Game ($F(1.84,91.77) = 3.35$, $\eta^{2}  = 0.04$,  $p = .043$, Greenhouse-Geisser correction to degrees of freedom applied to correct for non-sphericity).  We then run statistical tests on early round scores by game and opponent against the null hypothesis of 0 (no transfer). For level-1 facing players, there is evidence of learning transfer from RPS to both FWG ($t(150) = 3.96$, $p < .001$) and Numbers ($t(150) = 3.74$, $p < .001$). For level-2 facing players, there is evidence for transfer from RPS to the similar game FWG, albeit scores are lower than for level-1 player ($t(150) = 2.48$, $p = .01$) but not to the dissimilar game of Numbers.

### Discussion

The results of Experiment 1 indicate that participants were able to learn successful strategies which exploited the deviation from Nash-optimal play of their opponents. Moreover, they were able to transfer knowledge about their opponent to other games, but this was moderated by the type of game and opponent. There was evidence of transfer from the RPS game to the similar FWG game for both opponents. However, transfer to the dissimilar Numbers game was only evident for participants facing a level-1 opponent. <!-- These results indicate that learning transfer to the more dissimilar game (NUMBERS) we found earlier is exclusively driven by level-1 facing players, as average early round scores in the NUMBERS game of level-2 facing players are close to 0. Therefore, both participants facing level-1 and level-2 agents can transfer learning to the similar FWG game, but only those facing the less sophisticated opponent are able to generalise to the less similar NUMBERS game. -->

@@

In Experiment 2, we aimed to obtain a stronger test of learning transfer. <!-- We ran a second experiment with various differentiated features to improve the opportunity to measure learning transfer.--> Instead of facing a single level-1 or level-2 opponent throughout all games, participants face both types of opponents sequentially in this experiment. To perform well against both opponents, participants would need to learn distinct strategies in each game. To reduce effects of increased memory load due to facing distinct opponents, we provided participants access to the history of play against an opponent within each game (see Figure \@ref(fig:feedback-rps-exp2)). Finally, we changed the third game to a penalty shootout game, with participants aiming to score a goal and opponents playing the role of goal keepers. Whilst this game has the same number of actions as the first two (aim left, center, or right), it is strategically dissimilar. Unlike the Numbers game in Experiment 1, the Shootout game does not have a cyclical hierarchy between actions, making it harder to win through a heuristic based on this cyclicity.

<!-- Because there were two distinct opponents, requiring potentially holding two opponent models in memory, we also made it easier to recall the results of past rounds by providing participants with the opportunity to see the history of the game since the beginning of each interaction. Figure\ \@ref(fig:feedback-rps-exp2) shows an example of showing interaction history in the RPS game. Finally, we changed the third game to a penalty shootout game, whith participants aiming to score a goal and opponents playing the role of goal keepers.  Whilst this game has the same number of actions as the first two, it is strategically dissimilar. Unlike the third game in the first experiment, it did not have a cyclical hierarchy between actions, making it harder to win by just following simple heuristics leveraging this cyclicality. If we see evidence for differential play against opponents, it would show participants adapting their strategies to the opponent they are facing, which is indicative of opponent modelling. -->

@@

A total of 48 participants (21 females, 28 males, 1 unknown) were recruited via the Prolific platform, none of which took part in Experiment 1. The average age was 30.2 years, and the mean duration to complete the task was 39 minutes. Participants were incentivised using a two-tier payment mechanism: a fixed fee of £2.5 for completing the experiment plus a performance linked bonus that averaged £1.32.

@@

knitr::include_graphics("images/shootout.jpg")
```

@@

-->
